Component response rate variation underlies the stability of complex systems

### 2020-03-13 notebook/ms.Rmd
Add first paragraph of methods for constructing random and structured networks, and start the second

<!--- Need a section here on how the structured networks were built --->

>>>>

**Construction of random and structured networks**. I used the R programming language for all numerical simulations and analyses [@Rproject]. Purely random networks were generated by sampling off-diagonal elements from an i.i.d $A_{ij} \sim \mathcal{N}(0, \sigma_{A})$ with a probability $C$ (unsampled elements were set to zero). Diagonal elements $A_{ii}$ were set to $-1$. Elements of $\gamma$ were simulated i.i.d. from a distribution with positive support (typically $\gamma_{i} \sim \mathcal{U}(0, 2)$). Random $\mathbf{A}$ matrices with correlated elements $A_{ij}$ and $A_{ji}$ were built using Cholesky decomposition. Competitor networks in which off-diagonal elements $A_{ij} \geq 0$ were constructed by first building a random $\mathbf{A}$, then flipping the sign of any elements in which $A_{ij} < 0$. Similarly, mutualist networks were constructed by building a random $\mathbf{A}$, then flipping the sign of elements where $A_{ij} > 0$. Predator-prey networks were constructed by first buing a random $\mathbf{A}$, then flipping the sign of either $A_{ij}$ or $A_{ji}$ if $A_{ij} \times A_{ji} > 0$.

Small-world networks were constructed using the method of Watts and Strogatz [@Watts1998]. A random matrix $\mathbf{J}$ was generated, then a network $\mathbf{S}$ to get $\mathbf{A} = \mathbf{J} \odot \mathbf{S}$

### 2020-03-13 notebook/ms.tex
Add first paragraph of methods for constructing random and structured networks, and start the second

@@

@@

\textbf{Figure 1: Eigenvalue distributions of random complex systems.}
Each panel shows the real (x-axis) and imaginary (y-axis) parts of
\(S =\) 400 eigenvalues from random \(S \times S\) matrices.
(\(\textbf{a}\)) A system represented by a matrix \(\mathbf{A}\), in
which all elements are sampled from a normal distribution with
\(\mu = 0\) and \(\sigma_{A} = 1/\sqrt{S}\). Points are uniformly
distributed within the blue circle centred at the origin with a radius
of \(\sigma_{A} \sqrt{S} = 1\). (\(\textbf{b}\)) The same system as
\(\textbf{a}\) after including variation in the response rates of \(S\)
components, represented by the diagonal matrix \(\gamma\), such that
\(\mathbf{M} = \gamma\mathbf{A}\). Elements of \(\gamma\) are randomly
sampled from a uniform distribution from \(\min = 0\) to \(\max = 2\).
Eigenvalues of \(\mathbf{M}\) are then distributed non-uniformly within
the red circle centred at the origin with a radius of
\(\sqrt{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})S} \approx\) 1.17.
(\(\textbf{c}\)) A different random system \(\mathbf{A}\) constructed
from the same parameters as in \(\textbf{a}\), except with diagonal
values of \(-1\). (\(\textbf{d}\)) The same system \(\textbf{c}\) after
including variation in component response rates, sampled from
\(\mathcal{U}(0, 2)\) as in \(\textbf{b}\).

>>>>

\textbf{Construction of random and structured networks}. I used the R
programming language for all numerical simulations and
analyses\textsuperscript{\protect\hyperlink{ref-Rproject}{31}}. Purely
random networks were generated by sampling off-diagonal elements from an
i.i.d \(A_{ij} \sim \mathcal{N}(0, \sigma_{A})\) with a probability
\(C\) (unsampled elements were set to zero). Diagonal elements
\(A_{ii}\) were set to \(-1\). Elements of \(\gamma\) were simulated
i.i.d. from a distribution with positive support (typically
\(\gamma_{i} \sim \mathcal{U}(0, 2)\)). Random \(\mathbf{A}\) matrices
with correlated elements \(A_{ij}\) and \(A_{ji}\) were built using
Cholesky decomposition. Competitor networks in which off-diagonal
elements \(A_{ij} \geq 0\) were constructed by first building a random
\(\mathbf{A}\), then flipping the sign of any elements in which
\(A_{ij} < 0\). Similarly, mutualist networks were constructed by
building a random \(\mathbf{A}\), then flipping the sign of elements
where \(A_{ij} > 0\). Predator-prey networks were constructed by first
buing a random \(\mathbf{A}\), then flipping the sign of either
\(A_{ij}\) or \(A_{ji}\) if \(A_{ij} \times A_{ji} > 0\).

Small-world networks were constructed using the method of Watts amd
Strogatz\textsuperscript{\protect\hyperlink{ref-Watts1998}{15}}.

@@

\hypertarget{ref-Rproject}{}
31. R Core Team. \emph{R: A language and environment for statistical
computing}. (R Foundation for Statistical Computing, 2018).

@@

\textbf{Figure 1: Eigenvalue distributions of random complex systems.}
Each panel shows the real (x-axis) and imaginary (y-axis) parts of
\(S =\) 400 eigenvalues from random \(S \times S\) matrices.
(\(\textbf{a}\)) A system represented by a matrix \(\mathbf{A}\), in
which all elements are sampled from a normal distribution with
\(\mu = 0\) and \(\sigma_{A} = 1/\sqrt{S}\). Points are uniformly
distributed within the blue circle centred at the origin with a radius
of \(\sigma_{A} \sqrt{S} = 1\). (\(\textbf{b}\)) The same system as
\(\textbf{a}\) after including variation in the response rates of \(S\)
components, represented by the diagonal matrix \(\gamma\), such that
\(\mathbf{M} = \gamma\mathbf{A}\). Elements of \(\gamma\) are randomly
sampled from a uniform distribution from \(\min = 0\) to \(\max = 2\).
Eigenvalues of \(\mathbf{M}\) are then distributed non-uniformly within
the red circle centred at the origin with a radius of
\(\sqrt{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})S} \approx\) 1.15.
(\(\textbf{c}\)) A different random system \(\mathbf{A}\) constructed
from the same parameters as in \(\textbf{a}\), except with diagonal
values of \(-1\). (\(\textbf{d}\)) The same system \(\textbf{c}\) after
including variation in component response rates, sampled from
\(\mathcal{U}(0, 2)\) as in \(\textbf{b}\).
